Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2016-03-24 RMS/submission.tex
finished related work

We have analyzed our model/implementation in ten different scenarios,
including both realistic ones provided by our industrial partner and
experimental ones designed by ourselves, four of them are described
below:
\begin{itemize}
\item Scenario~(i) with two tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=3ms$, $T_2=25ms$ and $C_2=7ms$.
\item Scenario~(ii) with two tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=2ms$, $T_2=25ms$ and $C_2=2.3ms$.
\item Scenario~(iii) with three tasks $\tau_1$, $\tau_2$ and $\tau_3$:
  $T_1=5ms$, $C_1=2.7ms$, $T_2=10ms$, $C_2=2ms$, $T_3=25ms$ and
  $C_3=3ms$.
\item Scenario~(iv) with three tasks $\tau_1$, $\tau_2$ and $\tau_3$:
  $T_1=5ms$, $C_1=2.5ms$, $T_2=10ms$, $C_2=1.5ms$, $T_3=15ms$ and
$C_3=4.5ms$.
\end{itemize}

@@

Considering schedulability test, Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73} gave the famous sufficient
condition that a set of periodic tasks is schedulable with respect to
RMS if $\displaystyle \Sigma^n_{i=1} C_i/T_i \le n(2^{1/n}-1)$
holds. Then a more sufficient condition, known as \emph{Hyperbolic
  Bound}, which has the same complexity as Liu and Layland's, was
proposed in~\cite{DBLP:journals/tc/BiniBB03}. On the other hand,
necessary and sufficient conditions for schedulability were derived
independently in~\cite{DBLP:journals/rts/SpruntSL89}
and~\cite{audsley1993deadline}, requiring more sophisticated analysis
on the task set. Nevertheless, all these results take no overhead into
account, being not as realistic as ours. Katcher et al. did consider
overhead in their schedulability analysis, under several models based
on different kinds of popular
implementations~\cite{DBLP:journals/tse/KatcherAS93}.  However, our
target implementation is not in their scope. Furthermore, compared
with those theoretical analyses, our approach based on formal modeling
and verification has three advantages. One is that if our
schedulability test answers ``no'', it returns at the same time a real
counterexample, which is able to guide our engineer to adjust the
design, by changing either the priorities or even the scheduling
algorithm. The second is that, when a fresh scheduling strategy is
applied, our analysis can be adjusted only by modifying the model,
while theoretical approaches may need thorough analyses and reasoning.
The last one is that, considering overhead and details of hardware do
introduce some kind of non-determinism into the model. For example, in
our model, if the running task is complete ``right'' at the time when
an interrupt request is performed, two different behaviors are
possible: (i) the system performs task switching, during which the
interrupt request is masked, hence the scheduling may be delayed; (ii)
the system answers the interrupt request immediately.

\cite{DBLP:conf/iceccs/CuiDT14} and~\cite{TianD2011} also made use of
model checking to analyze the RMS algorithm along the same line but
with different languages and tools. They considered only the ideal
setting as in~\cite{DBLP:journals/jacm/LiuL73}, instead of an
implementation which contains much more complex details. Our model of
the implementation easily degenerates to a model of the RMS algorithm,
if we let the times for scheduling and switching be zero.

>>>>

We have analyzed our model/implementation in ten different scenarios,
including both realistic ones provided by our industrial partner and
experimental ones designed by ourselves, four of them are described
below:
\begin{itemize}
\item Scenario~(i) with two tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=3ms$, $T_2=25ms$ and $C_2=7ms$.
\item Scenario~(ii) with two tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=2ms$, $T_2=25ms$ and $C_2=2.3ms$.
\item Scenario~(iii) with three tasks $\tau_1$, $\tau_2$ and $\tau_3$:
  $T_1=5ms$, $C_1=2.7ms$, $T_2=10ms$, $C_2=2ms$, $T_3=25ms$ and
  $C_3=3ms$.
\item Scenario~(iv) with three tasks $\tau_1$, $\tau_2$ and $\tau_3$:
  $T_1=5ms$, $C_1=2.5ms$, $T_2=10ms$, $C_2=1.5ms$, $T_3=15ms$ and
$C_3=4.5ms$.
\end{itemize}
Note that thanks to the powerful expressivity of Real-Time Maude, we
only need to define an initial state of sort \verb|System| to specify
a given task set. No necessity to modify the model is needed.

@@

Considering schedulability test, Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73} gave the famous sufficient
condition that a set of periodic tasks is schedulable with respect to
RMS if $\displaystyle \Sigma^n_{i=1} C_i/T_i \le n(2^{1/n}-1)$
holds. Then a more sufficient condition, known as \emph{Hyperbolic
  Bound}, which has the same complexity as Liu and Layland's, was
proposed in~\cite{DBLP:journals/tc/BiniBB03}. On the other hand,
necessary and sufficient conditions for schedulability were derived
independently in~\cite{DBLP:journals/rts/SpruntSL89}
and~\cite{audsley1993deadline}, requiring more sophisticated analysis
on the task set. Nevertheless, all these results take no overhead into
account, being not as realistic as ours. Katcher et al. did consider
overhead in their schedulability analysis, under several models based
on different kinds of popular
implementations~\cite{DBLP:journals/tse/KatcherAS93}.  However, our
target implementation is not in their scope. Furthermore, compared
with those theoretical analyses, our approach based on formal modeling
and verification has three advantages. One is that if our
schedulability test answers ``no'', it returns at the same time a real
counterexample, which is able to guide our engineer to adjust the
design, by changing either the priorities or even the scheduling
algorithm. The second is that, when a fresh scheduling strategy is
applied, our analysis can be adjusted only by modifying the model,
while theoretical approaches may need thorough analyses and reasoning.
The last one is that, considering overhead and details of hardware do
introduce some kind of non-determinism into the model. For example, in
our model, if the running task is complete \emph{right} at the time
when an interrupt request is performed, two different behaviors are
possible: (i) the system performs task switching, during which the
interrupt request is masked, hence the scheduling may be delayed; (ii)
the system answers the interrupt request immediately, the switching
being delayed. Tackling such non-determinism via theoretical analyses
seems more complicated than our automatic approach.

\cite{TianD2011} and~\cite{DBLP:conf/iceccs/CuiDT14} also made use of
model checking to analyze the RMS algorithm along the same line but
with different languages and tools. The RMS algorithm is analyzed
using an extension of SPIN~\cite{DBLP:journals/tse/Holzmann97}
in~\cite{TianD2011}, while the logic programming language
TMSVL~\cite{DBLP:conf/icfem/HanDW12} and its model checker are applied
in~\cite{DBLP:conf/iceccs/CuiDT14}. The work presented
in~\cite{TianD2011} and~\cite{DBLP:conf/iceccs/CuiDT14} has two main
differences with ours. The first one, which is also the motivational
one, is that they considered only the ideal setting of the algorithm
as in~\cite{DBLP:journals/jacm/LiuL73}, instead of an implementation
which contains much more complex details. In fact, our model of the
implementation can easily degenerate to a model of the RMS algorithm,
if we let the times for scheduling and switching be zero. The other
difference is that when adding a new task into the task set, the
models in~\cite{TianD2011} and~\cite{DBLP:conf/iceccs/CuiDT14} should
be modified by explicitly defining a sub-model of the new task and its
behaviors. In particular, the scheduling part of the model
in~\cite{TianD2011} needs adjustments as well to include a new
task. Nevertheless, a new task set is specified in our approach merely
by giving a new initial state, without necessity to modify the model,
as already emphasized in Section~\ref{ss:results}. On the other hand,
\cite{TianD2011} used a discrete time domain in the model, while we
employ a generic time domain, which is flexible to be instantiated on
either discrete or dense ones. In~\cite{DBLP:conf/iceccs/CuiDT14}, the
time domain is dense, and a modeling strategy like our maximal time
sampling strategy is applied to reduce the state space. However, a
completeness statement like Theorem~\ref{t:main} was not given
in~\cite{DBLP:conf/iceccs/CuiDT14}.
